Linting Linearity in Core/System FC

### 2023-01-25 report.tex
optimizations subsections list
Tweaks

Linearity in Haskell stands out % from linearity in Rust and Idris2
two distinct reasons:

@@

% % TODO: Revise the story above instead of dumping so much information

@@

Haskel's intermediate representation that is statically typed

\begin{itemize}
    \item What is Core (IR) e as particularidades
    \item Em cima do SystemFC
    \item Optimizações feitas no Core
    \item Base SystemFC tem de ser "merged" com o linear lambda calculus
    \item Ou porque é que é importante o Core ter linearidade também/a que nível
        foi introduzido
    \item Falar das coercions em especifico
    \item E das local equalities que são implicitas no front end
\end{itemize}

@@

@@

>>>>

Linearity in Haskell stands out from linearity in Rust and Idris2
two distinct reasons:

@@

@@

Haskell is a big and expressive language with a lot of syntatic constructs and
many, many features. However, the whole of Haskell can be desugared down to a
very simple and elegant, explicitly typed, intermediate representation called
\textbf{Core}. To illustrate the difference in complexity, in GHC, the source
language (the big and complex Haskell) is defined through dozens of datatypes
and hundreds of constructors, while Core is defined in 3 types (expressions,
types, and coercions) and 15 constructors~\cite{}.
%
Core is unique as a strongly-typed, lazy, purely functional intermediate
language akin to a polymorphic lambda calculus, with GHC as the only
production-ready compiler having such a typed intermediate representation.

Core allows us to reason about the entirety of Haskell in a much smaller and
elegant functional language -- analysis, optimizating transformations, and code
generation are all done on Core, not Haskell.
%
Of equal importance is Core being explicitly typed, which makes Core an internal
consistency tool that validates desugaring and optimization passes. Desugaring
and optimizing transformations must produce well-typed Core, and the Core
typechecker can guarantee that Core programs are well-typed, thus providing a
verification layer of the correctness of the optimizing transformations, of the
desugarer, and of their implementations.
%
Finally, the expressive but simple type-system in Core serves as a sanity-check
for all the extensions to the source language type system -- if we can desugar a
type system feature into Core then the feature must be sound by reducibility.
Effectively, any feature added to Haskell is only syntatic sugar if it can be
desugared to Core.

Core is typed very differently from Haskell because Core is explicitly typed and
its type system is based on the $System~F_C$~\cite{systemfc} type system, while
Haskell is implicitly typed and its type system is based on the constraint-based
type inference system $OutsideIn(X)$~\cite{outsideinx}. Therefore, typechecking
Core is simple and fast and no inference must be done and, on the contrary,
Haskell's typechecker is much more complicated and requires ad-hoc restrictions
to make type inference feasible~\cite{}.
%
Haskell is typechecked in addition to Core to (i) do type inference and (ii)
provide meaningful type errors. If Haskell wasn't typechecked as it is and
instead we only typechecked Core, everything (e.g.  all binders) would have to
be explicitly typed and all error messages would refer to the intermediate
language rather than the written program, which is known to be undesirable.

\begin{itemize}
    % \item What is Core (IR) e as particularidades
    \item Em cima do SystemFC
    \item Optimizações feitas no Core
    \item Base SystemFC tem de ser "merged" com o linear lambda calculus
    \item Ou porque é que é importante o Core ter linearidade também/a que nível
        foi introduzido
    \item Falar das coercions em especifico
    \item E das local equalities que são implicitas no front end
\end{itemize}

@@

\section{Core-To-Core Optimizations}

% like inlining, its inverse (CSE), beta-reduction, lambda-lifting, eta-reduction/eta-expansion, binder-swap, case-of-case, case-of-know-constructor, float-out, float-in, worker/wrapper split (this one is big, in comparison), etc…)

\subsection{Inlining}

\subsection{Common Sub-expression elimination}

\subsection{Beta-reduction}
\subsection{Lambda lifting}
\subsection{Eta-reduction/Eta-expansion}

\subsection{Binder-swap}

\subsection{Case-of-case}
\subsection{Case-of-known-constructor}
\subsection{Float-out and Float-in}
\subsection{Worker/wrapper split}

@@

% TODO: - In the letrec case: congratulations on finding an inference algorithm
% (though I will want to see the proof some day, I don't see why it works yet).
% You should clarify that, during linting, you will have a usage environment
% annotation and won't need to run the inference algorithm. This algorithm is
% only needed when you first create a letrec.
